---
title: "Descriptive Statistics for the VQI FBVAR Dataset"
author: "Jennifer Ci, Thu Vu, Lily Hanyi Wang"
output: pdf_document
---


```{r library, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,fig.width = 10)
library(tidyverse)
library(table1)
library(Hmisc)
library(ggplot2)
library(ggpubr)
```

```{r setup wd}
## ------------- working directories for Lily ----------
wd_lily = '/Users/hanyiwang/Desktop/Comparative-analysis-of-treatments-of-CAA'
path_lily = c("../data/FBVAR.csv")


## ------------- working directories for Jenn ----------
#wd_jenn = '/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA'
#path_jenn = c(
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210712/TEV#AR_International_LTF_r12_2_14_20210701.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210712/TEV#AR_International_PROC_r12_2_14_20210701.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210901/TEVA#R_International_LTF_r12_2_14_20210901.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210901/TEV#AR_International_PROC_r12_2_14_20210901.csv")

## ------------- read data ----------
setwd(wd_lily)
FBVAR = read.csv(path_lily)

#setwd(wd_jenn)
#TEVAR_LTF_07 = read.csv(path_jenn[1])
#TEVAR_PROC_07 = read.csv(path_jenn[2])
#TEVAR_LTF_09 = read.csv(path_jenn[3])
#TEVAR_PROC_09 = read.csv(path_jenn[4])
```


## population of interest: the asymptomatic and symptomatics groups. 

```{r}
## ------------- population of interest ----------
table1(~  PRESENTATION, data = FBVAR)
```

## Patient demographic and co-morbidities

**Table: A comparison of the baseline demographic and co-morbidities characteristics for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: Patient demographic and co-morbidities----------

table1(~ AGE+AGECAT+GENDER+ETHNICITY+ RACE+TRANSFER+PRIMARYINSURER+ 
         LIVINGSTATUS+ PREOP_FUNCSTATUS+PRIOR_CVD +PRIOR_CAD+PRIOR_CHF+COPD+
         DIABETES+PREOP_DIALYSIS+HTN+ PREOP_SMOKING+PRIOR_CABG+PRIOR_PCI+PRIOR_ANEURREP+
         STRESS+PREOP_CREAT+DC_ASA+DC_P2Y+DC_STATIN+PREOP_CREAT
         | PRESENTATION, data = FBVAR,overall="Total")
```


## Operative Variables

**Table: A comparison of the operative characteristics for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table:  Operative Variables----------
table1(~ PRIOR_AORSURG+
         PATHOLOGY+PREOP_MAXAAADIA+URGENCY+
         PATHOLOGY_ANEURYSM_TYPE+PATHOLOGY_DISSECT_TYPE+PATHOLOGY_DISSECT_ONSET_DAYS+
         GENE+PROXZONE_DISEASE+DISTZONE_DISEASE+ extent+
         ANESTHESIA+CONTRAST+EBL+FLUOROTIME+INTRAOP_PRBC+TOTALPROCTIME+IVUSTEE+
         ACCESS+ #ACCESS_R+ACCESS_L+
         #LRGST_SHEATH_SIZE_R+LRGST_SHEATH_SIZE_L+
         ARMNECK_ACCESS+ #ARMNECK_ACCESS_LOC+
         AORDEV_NUM+AORDEV_CMOD+# STAGEDAORTRT+
         DEV_GTYPE+ # DEV1_GTYPE+DEV2_GTYPE+DEV3_GTYPE+
         ILIACDEV_END_R+ILIACDEV_END_L+
         BRANCH_STAGED+BRANCH_LSUB+BRANCH_CELIAC+BRANCH_SMA+BRANCH_RRENAL+BRANCH_LRENAL+
         ANESTHESIA_GEN_TIMEEXT+POSTOP_SPINALDRAIN+
         R_CELIAC+R_DISTATTZONE+R_GDPROXIMAL+R_GRFTCONFIG+
         R_LT_RENAL+R_PRATTZONE+R_RT_RENAL+R_SMA+
         lrenal+rrenal+sma+celiac+lsub
       | PRESENTATION, data = FBVAR)

FBVAR %>% filter(lrenal=='Chimney'|rrenal=='Chimney'|sma=='Chimney'|celiac=='Chimney'|lsub=='Chimney') %>% select(lrenal,rrenal,sma,celiac)

```

The levels of `lrenal`,`rrenal`,`sma`,`celiac`,`lsub` are really messy.

`r nrow(FBVAR %>% filter(lrenal=='Chimney'|rrenal=='Chimney'|sma=='Chimney'|celiac=='Chimney'|lsub=='Chimney') )` patients have at least one 'Chimney'.

`r nrow(FBVAR %>% filter(lrenal=='Occluded/Covered'|rrenal=='Occluded/Covered'|sma=='Occluded/Covered'|celiac=='Occluded/Covered'|lsub=='Occluded/Covered') )` patients have at least one 'Occluded/Covered'.

`r nrow(FBVAR %>% filter(lrenal=='Scallop/Fen/Branch'|rrenal=='Scallop/Fen/Branch'|sma=='Scallop/Fen/Branch'|celiac=='Scallop/Fen/Branch'|lsub=='Scallop/Fen/Branch') )` patients have at least one 'Scallop/Fen/Branch'.

`r nrow(FBVAR %>% filter(lrenal=='None'& rrenal=='None'& sma=='None'& celiac=='None'& lsub=='None') )` patients have all 'None'.


## Outcomes

**Table 3: A comparison of the long term follow-up outcomes for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: primary outcomes----------
table1(~ DEAD+PROC_SURVIVALDAYS | PRESENTATION, data = FBVAR)
```

**Table 3: A comparison of the procedure outcomes for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: secondary outcomes----------
table1(~ TOTAL_LOS+POSTOP_LOS+
         AORDEV_TECHSUCC+CONVTOOPEN+R_ENDOLEAK_AT_COMPLETION+
         #BRANCH_LSUB_POST+BRANCH_CELIAC_POST+BRANCH_SMA_POST+BRANCH_RRENAL_POST+
         #BRANCH_LRENAL_POST+BRANCH_RCOMILI_POST+BRANCH_LCOMILI_POST+
         BRANCH_POST+ICUSTAY+
         POSTOP_PRBC+POSTOP_VASO+POSTOP_HIGHCREAT+POSTOP_COMPLICATIONS+
         R_POSTOP_HEMATOMA+R_POSTOP_SITEOCC+
         POSTOP_AH+#POSTOP_MI+POSTOP_DYSRHYTHMIA+POSTOP_CHF+
         POSTOP_CEREBROSX+POSTOP_RESPIRATORY+
         POSTOP_DIALYSIS+POSTOP_ARMEMBO+POSTOP_LEGEMBO+POSTOP_LEGCOMPART+POSTOP_INTISCH+
         R_POSTOP_BOWELISCH+POSTOP_RENALISCH+R_POSTOP_SSI+POSTOP_SPINAL_ISCHEMIA+
         RETX_R_RTOR+#RETX+
         DC_STATUS+R_LE_ISCH+R_POSTOP_RENAL#+R_RTOR
       | PRESENTATION, data = FBVAR)

## ------------- Survival curves----------

```

## Volume Variables 
  
Volume Variables: `REGIONID`, `CENTERID`, `PHYSICIANID`

```{r}
## ------------- clustering variables----------

#FBVAR %>% select(REGIONID) %>% table() 
#FBVAR %>% select(CENTERID) %>% table() 
#FBVAR %>% select(PHYSICIANID) %>% table() 

## ------------- plots of volume----------
center_vol = as.data.frame(FBVAR %>% select(CENTERID) %>% table())
phys_vol = as.data.frame(FBVAR %>% select(PHYSICIANID) %>% table())

p1 = ggplot(data = center_vol, aes(x=CENTERID, y=Freq)) + 
  geom_point() +
  labs(title = 'Volume of centers',x='Center ID',y='Volume')

p2 = ggplot(data = center_vol, aes(x='', y=Freq)) + 
  geom_boxplot() +
  labs(title = 'Boxplot of center volumes',x='',y='Volume')

print(ggarrange(p1, p2, widths = c(20,10),ncol = 2, nrow = 1, align = "h"))

p3 = ggplot(data = phys_vol, aes(x=PHYSICIANID, y=Freq)) + 
  geom_point() +
  labs(title = 'Volume of physicians',x='Physician ID',y='Volume')

p4 = ggplot(data = phys_vol, aes(x='', y=Freq)) + 
  geom_boxplot() +
  labs(title = 'Boxplot of physicians volumes',x='',y='Volume')

print(ggarrange(p3, p4, widths = c(20,10),ncol = 2, nrow = 1, align = "h"))

```

`r nlevels(factor(FBVAR$REGIONID))` regions, `r nlevels(factor(FBVAR$CENTERID))` centers, `r nlevels(factor(FBVAR$PHYSICIANID))` physicians. 

Quantiles of centers' volume: `r quantile(center_vol[order(center_vol$Freq),]$Freq)`

Quantiles of physicians' volume: `r quantile(phys_vol[order(phys_vol$Freq),]$Freq)`


\newpage
## Code Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,include=TRUE}
```


