---
title: "preprocessing"
author: "Jennifer Ci, Thu Vu, Lily Hanyi Wang"
date: "2022-10-26"
output: pdf_document
---

```{r library, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
library(tidyverse)
library(table1)

```


```{r setup wd}
## ------------- working directories for Lily ----------
wd_lily = '/Users/hanyiwang/Desktop/Comparative-analysis-of-treatments-of-CAA'
path_lily = c(
  "../data/TEVAR_International_20210712/TEVAR_International_LTF_r12_2_14_20210701.csv",
  "../data/TEVAR_International_20210712/TEVAR_International_PROC_r12_2_14_20210701.csv",
  "../data/TEVAR_International_20210901/TEVAR_International_LTF_r12_2_14_20210901.csv",
  "../data/TEVAR_International_20210901/TEVAR_International_PROC_r12_2_14_20210901.csv")


## ------------- read data ----------
setwd(wd_lily)
TEVAR_LTF_07 = read.csv(path_lily[1])
TEVAR_PROC_07 = read.csv(path_lily[2])
#TEVAR_LTF_09 = read.csv(path_lily[3])
#TEVAR_PROC_09 = read.csv(path_lily[4])
```

## Dataset `TEVAR_PROC_07` to start with.

```{r}
## ------------- data cleaning----------
TEVAR_PROC_07 = TEVAR_PROC_07 %>% 
  mutate(DEAD=factor(DEAD)) %>%
  mutate(PRESENTATION = factor(PRESENTATION,levels = c(0,1,2),
                               labels = c('Asymptomatic','Symptomatic','Rupture'))) %>%
  mutate(AGECAT = factor(AGECAT,levels = c(1,2,3,4,5,6,7),
                         labels = c('<40','40-49','50-59','60-69','70-79','80-89','>89'))) %>%
  mutate(GENDER=factor(GENDER,levels=c(1,2),
                       labels=c('male','female')))

```


## population of interest: the asymptomatic and symptomatics groups. 

`PRESENTATION`: 0 = Asymptomatic,1 = Symptomatic,2 = Rupture

Is this the correct variable? 

What does rupture mean? 

Should we just delete data missing symptom info?

```{r}
## ------------- population of interest ----------
table1(~  PRESENTATION, data = TEVAR_PROC_07)
```

## potential outcome variables

Are these possible outcomes?

`TOTAL_LOS`: Length of Stay in days calculated by DISCHARGE_DT - ADMIT_DT \\

`POSTOP_LOS`: Length of Stay in days calculated by DISCHARGE_DT - SURGERY_DT

```{r}
## ------------- descriptive statistics table for outcomes----------
table1(~ factor(DEAD) + PROC_SURVIVALDAYS | PRESENTATION, data = TEVAR_PROC_07)
```

## patient condition variables:

`AGECAT`: 1 = <40,2 = 40-49,3 = 50-59,4 = 60-69,5 = 70-79,6 = 80-89,7 = >89

`GENDER`: 1 = Male,2 = Female

`PRIOR_CVD`: 0 = None,1 = hx stroke, asymptomatic,2 = hx stroke, minor deficit,3 = hx stroke, major deficit

`PRIOR_CAD`: 0 = None,1 = hx MI but no sx,2 = Stable angina,3 = Unstable angina or MI < 6 mos (retired since 09/12/2012),4 = MI < 6 mos,5 = Unstable angina

`PRIOR_CHF`: 0 = None,1 = Asymp, hx CHF,2 = Mild,3 = Moderate,4 = Severe

`COPD`: 0 = No,1 = Not Treated,2 = On Meds,3 = On Home Oxygen

`DIABETES`: 0 = None,1 = Diet,2 = Non-insulin Meds,3 = Insulin

`HTN`: History of hypertension; 0 = No, 1 = Yes (>=140/90 or history) (retired since 11/15/2016), 2 = Yes, controlled  [added on 04/13/2020] , 3 = Yes, uncontrolled  [added on 04/13/2020]

`PREOP_SMOKING`: 0 = Never,1 = Prior,2 = Current

`PRIOR_AORSURG`: Any aortic procedures performed on a separate date prior to the index procedure; 0 = None,1 = Open,2 = Endo,3 = Both,4 = Other (retired since 09/30/2014)


`PATHOLOGY`: 1 = Aneurysm,2 = Dissection,3 = Aneurysm from dissection,4 = Trauma,5 = Penetrating Ulcer (PAU),6 = Intramural Hematoma (IMH),7 = PAU with IMH,8 = Aortic Thrombus,9 = Other (Retired) (retired since 09/30/2014),10 = Aorto-esophageal Fistula (Retired) (retired since 09/30/2014),11 = Aorto-bronchial Fistula (Retired) (retired since 09/30/2014).

*Type of TAA?*

`PREOP_MAXAAADIA`: "Maximum total aortic diameter within the last 6 months (including true and false lumen in dissection) within diseased segment being treated, measured at right angle to centerline, ideally with 3-D imaging software. Min/max range: 16 to 190 mm."

*surgery year?*

```{r}
## ------------- descriptive statistics table for patients conditions----------
table1(~ GENDER+AGE+AGECAT+factor(PREOP_SMOKING)+factor(PRIOR_CVD)+factor(PRIOR_CAD)+
         factor(PRIOR_CHF)+factor(COPD)+factor(DIABETES)+factor(HTN)+factor(PRIOR_AORSURG)
       | PRESENTATION, data = TEVAR_PROC_07)

```


## Medical center info:

```{r}
## ------------- descriptive statistics table for other variables of interest----------

#TEVAR_PROC_07 %>% select(REGIONID) %>% table() 
#TEVAR_PROC_07 %>% select(CENTERID) %>% table() 
#TEVAR_PROC_07 %>% select(PHYSICIANID) %>% table() 
```

`r nlevels(factor(TEVAR_PROC_07$REGIONID))` regions, `r nlevels(factor(TEVAR_PROC_07$CENTERID))` centers, `r nlevels(factor(TEVAR_PROC_07$PHYSICIANID))` physicians. 

Most physicians only performed 1 or 2 procedures. Several performed over 100 procedures? Is that real? 

There are regions and centers that perfomed many procedures. Would that affect the outcome?

\newpage
### Code Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,include=TRUE}
```


