---
title: "preprocessing"
author: "Jennifer Ci, Thu Vu, Lily Hanyi Wang"
date: "2022-10-26"
output: pdf_document
---

```{r library, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
library(tidyverse)
library(table1)

```


```{r setup wd}
## ------------- working directories for Lily ----------
wd_lily = '/Users/hanyiwang/Desktop/Comparative-analysis-of-treatments-of-CAA'
path_lily = c(
  "../data/TEVAR_International_20210712/TEVAR_International_LTF_r12_2_14_20210701.csv",
  "../data/TEVAR_International_20210712/TEVAR_International_PROC_r12_2_14_20210701.csv",
  "../data/TEVAR_International_20210901/TEVAR_International_LTF_r12_2_14_20210901.csv",
  "../data/TEVAR_International_20210901/TEVAR_International_PROC_r12_2_14_20210901.csv")


## ------------- read data ----------
setwd(wd_lily)
TEVAR_LTF_07 = read.csv(path_lily[1])
TEVAR_PROC_07 = read.csv(path_lily[2])
#TEVAR_LTF_09 = read.csv(path_lily[3])
#TEVAR_PROC_09 = read.csv(path_lily[4])
```

## merge the datasets

## data cleaning based on inclusion, exclusion criteria

Exclusion criteria:

- exclude rupture groups.

- exclude pathology groups

```{r}
## ------------- data cleaning----------
TEVAR_PROC_07 = TEVAR_PROC_07 %>% 
  mutate(DEAD=factor(DEAD)) %>%
  mutate(PRESENTATION = factor(PRESENTATION,levels = c(0,1,2),
                               labels = c('Asymptomatic','Symptomatic','Rupture'))) %>%
  mutate(AGECAT = factor(AGECAT,levels = c(1,2,3,4,5,6,7),
                         labels = c('<40','40-49','50-59','60-69','70-79','80-89','>89'))) %>%
  mutate(GENDER=factor(GENDER,levels=c(1,2),
                       labels=c('male','female'))) %>%
  mutate(SURGYEAR=factor(SURGYEAR)) %>%
  mutate(PATHOLOGY=factor(PATHOLOGY)) 

## ------------- inclusion and exclusion----------

```


## population of interest: the asymptomatic and symptomatics groups. 

`PRESENTATION`: 0 = Asymptomatic,1 = Symptomatic,2 = Rupture


```{r}
## ------------- population of interest ----------
table1(~  PRESENTATION, data = TEVAR_PROC_07)
```



## Demographic history

GENDER, BMI(USING HEIGHT AND WEIGHT)?, R_PREOP_AMBUL(Preop ambulatory status), AGE (Patient age at procedure), AGECAT, PATHOLOGY (exclude 4, 8-11) 

`AGECAT`: 1 = <40,2 = 40-49,3 = 50-59,4 = 60-69,5 = 70-79,6 = 80-89,7 = >89

`GENDER`: 1 = Male,2 = Female

```{r}
## ------------- table1: demographic----------
table1(~ GENDER+AGE+AGECAT
         | PRESENTATION, data = TEVAR_PROC_07)
```

## patient condition variables:

`PATHOLOGY`: 1 = Aneurysm,2 = Dissection,3 = Aneurysm from dissection,4 = Trauma,5 = Penetrating Ulcer (PAU),6 = Intramural Hematoma (IMH),7 = PAU with IMH,8 = Aortic Thrombus,9 = Other (Retired) (retired since 09/30/2014),10 = Aorto-esophageal Fistula (Retired) (retired since 09/30/2014),11 = Aorto-bronchial Fistula (Retired) (retired since 09/30/2014)

Old variable, used to be called Indication, now mapped to Pathology. Mapping detail: 1 = TAA->1 = Aneurysm; 2 = TAAA->1 = Aneurysm; 4 = Dissection->2 =  Dissection; 3=Aneurysm from dissection; 3 = Trauma->4=Trauma; 5 = Penetrating Ulcer->5 = Penetrating Ulcer (PAU); 6 = Aortic Intramural Hematoma->6 =  Intramural Hematoma (IMH); 7=PAU with IMH; 8=Aortic Thrombus

`PRIOR_CVD`: 0 = None,1 = hx stroke, asymptomatic,2 = hx stroke, minor deficit,3 = hx stroke, major deficit

`PRIOR_CAD`: 0 = None,1 = hx MI but no sx,2 = Stable angina,3 = Unstable angina or MI < 6 mos (retired since 09/12/2012),4 = MI < 6 mos,5 = Unstable angina

`PRIOR_CHF`: 0 = None,1 = Asymp, hx CHF,2 = Mild,3 = Moderate,4 = Severe

`COPD`: 0 = No,1 = Not Treated,2 = On Meds,3 = On Home Oxygen

`DIABETES`: 0 = None,1 = Diet,2 = Non-insulin Meds,3 = Insulin

`HTN`: History of hypertension; 0 = No, 1 = Yes (>=140/90 or history) (retired since 11/15/2016), 2 = Yes, controlled  [added on 04/13/2020] , 3 = Yes, uncontrolled  [added on 04/13/2020]

`PREOP_SMOKING`: 0 = Never,1 = Prior,2 = Current

`PRIOR_AORSURG`: Any aortic procedures performed on a separate date prior to the index procedure; 0 = None,1 = Open,2 = Endo,3 = Both,4 = Other (retired since 09/30/2014)

```{r}
## ------------- table2: anatomy ----------
table1(~ factor(PATHOLOGY)+
         factor(R_PREOP_AMBUL)+factor(PREOP_SMOKING)+factor(PRIOR_CVD)+
         factor(PRIOR_CAD)+factor(PRIOR_CHF)+factor(COPD)+factor(DIABETES)+factor(HTN)+
         factor(PRIOR_AORSURG)
       | PRESENTATION, data = TEVAR_PROC_07)


```

## other variables

Surgery year would affect outcome, since surgeons got more familiar with the surgery.

```{r}
plot(DEAD~SURGYEAR,data=TEVAR_PROC_07)
```


## Outcome variables

`PROC_SURVIVALDAYS`: This should be the longest known time of survival data available for the patient. 
Survival days are calculated as the Last Date of Contact (or Date of Death) for the patient - Procedure date for a procedure. Please refer to included Death and Survival Days Logic.pdf for additional details."

`POSTOP_LOS`: Length of Stay in days calculated by DISCHARGE_DT - SURGERY_DT


```{r}
## ------------- table3: outcomes----------
table1(~ factor(DEAD) + PROC_SURVIVALDAYS+POSTOP_LOS | PRESENTATION, data = TEVAR_PROC_07)

## ------------- Survival curves----------

```

## Clustering variables:

```{r}
## ------------- clustering variables----------

#TEVAR_PROC_07 %>% select(REGIONID) %>% table() 
#TEVAR_PROC_07 %>% select(CENTERID) %>% table() 
#TEVAR_PROC_07 %>% select(PHYSICIANID) %>% table() 
```

`r nlevels(factor(TEVAR_PROC_07$REGIONID))` regions, `r nlevels(factor(TEVAR_PROC_07$CENTERID))` centers, `r nlevels(factor(TEVAR_PROC_07$PHYSICIANID))` physicians. 

Most physicians only performed 1 or 2 procedures. Several performed over 100 procedures. Since the more surgeries a surgeon did, the more familiar he or she is. So we need to cluster on this.

*Cluster on centers and physicians*

\newpage
### Code Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,include=TRUE}
```


