---
title: "Univariate Analysis for the VQI FBVAR Dataset"
author: "Jennifer Ci, Thu Vu, Lily Hanyi Wang"
output: pdf_document
---


```{r library, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,fig.width = 10)
library(tidyverse)
library(table1)
library(survival)
library(Hmisc)
library(ggplot2)
library(ggpubr)

```

```{r setup wd}
## ------------- working directories for Lily ----------
wd_lily = '/Users/hanyiwang/Desktop/Comparative-analysis-of-treatments-of-CAA'
path_lily = c("../data/FBVAR.csv")


## ------------- working directories for Jenn ----------
#wd_jenn = '/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA'
#path_jenn = c(
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210712/TEV#AR_International_LTF_r12_2_14_20210701.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210712/TEV#AR_International_PROC_r12_2_14_20210701.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210901/TEVA#R_International_LTF_r12_2_14_20210901.csv",
#  "/Users/jenniferci/Desktop/Comparative-analysis-of-treatments-of-CAA/TEVAR_International_20210901/TEV#AR_International_PROC_r12_2_14_20210901.csv")

## ------------- read data ----------
setwd(wd_lily)
FBVAR = read.csv(path_lily)

#setwd(wd_jenn)
#TEVAR_LTF_07 = read.csv(path_jenn[1])
#TEVAR_PROC_07 = read.csv(path_jenn[2])
#TEVAR_LTF_09 = read.csv(path_jenn[3])
#TEVAR_PROC_09 = read.csv(path_jenn[4])
```

## p-value

We use Welch's Two Sample t-test for continuous variables and Pearson's Chi-squared Test for categorical variables.

```{r}
pvalue <- function(x, ...) {
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, Welch's Two Sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, Pearson's Chi-squared Test
        p <- chisq.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

```

## Descriptive statistics tables


### population of interest 

```{r}
## ------------- population of interest ----------
table1_POI = table1(~  PRESENTATION, data = FBVAR)
knitr::kable(table1_POI)
```

### Patient demographic and co-morbidities

**Table: A comparison of the baseline demographic and co-morbidities characteristics for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: Patient demographic and co-morbidities----------

table1_CMB = table1(~   AGE+AGECAT+GENDER+ETHNICITY+ RACE+ TRANSFER+ PRIMARYINSURER+ LIVINGSTATUS+ PREOP_FUNCSTATUS+ PRIOR_CVD+ PRIOR_CAD+ PRIOR_CHF+ COPD+ DIABETES+ PREOP_DIALYSIS+ HTN+ PREOP_SMOKING+ PRIOR_CABG+ PRIOR_PCI+ PRIOR_ANEURREP+ STRESS+ PREOP_CREAT+ DC_ASA+DC_P2Y+DC_STATIN
       | PRESENTATION, data = FBVAR,overall=F, extra.col=list(`P-value`=pvalue))
knitr::kable(table1_CMB)

```


### Operative Variables

**Table: A comparison of the operative characteristics for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table:  Operative Variables----------
table1_OPR = table1(~ PRIOR_AORSURG+ PATHOLOGY+ PREOP_MAXAAADIA+ URGENCY+ PATHOLOGY_ANEURYSM_TYPE+ PATHOLOGY_DISSECT_TYPE+ PROXZONE_DISEASE+ GENHIST+ DISTZONE_DISEASE+ extent+ ANESTHESIA+ CONTRAST+ EBL+FLUOROTIME+ INTRAOP_PRBC+TOTALPROCTIME+IVUSTEE+ ACCESS+ARMNECK_ACCESS+ AORDEV_NUM+ AORDEV_CMOD+ DEV_GTYPE+ ILIACDEV_END_R+ ILIACDEV_END_L+ BRANCH_STAGED+ BRANCH_LSUB+ BRANCH_CELIAC+ BRANCH_SMA+BRANCH_RRENAL+ BRANCH_LRENAL+ ANESTHESIA_GEN_TIMEEXT+ POSTOP_SPINALDRAIN+lrenal+rrenal+sma+celiac+lsub+factor(NUM_TREATED_BRANCHES)+ factor(NUM_TREATED_RENALS)+ OCCLUDED_RENAL+ OCCLUDED_SMA+ OCCLUDED_CELIAC
       | PRESENTATION, data = FBVAR,overall=F, extra.col=list(`P-value`=pvalue))

knitr::kable(table1_OPR)

```


### Outcomes

**Table 3: A comparison of the long term follow-up outcomes for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: primary outcomes----------
table1_POC = table1(~ DEAD+PROC_SURVIVALDAYS+LTF_NUM_REINT | PRESENTATION, data = FBVAR,overall=F, extra.col=list(`P-value`=pvalue))

knitr::kable(table1_POC)

```

**Table 3: A comparison of the procedure outcomes for symptomatic versus asymptomatic patients who undergo the F-BEVAR procedure**

```{r}
## ------------- table: secondary outcomes----------
table1_SOC = table1(~ TOTAL_LOS+ POSTOP_LOS+ AORDEV_TECHSUCC+ CONVTOOPEN+ LEAKATCOMP_NONE+ ICUSTAY+ POSTOP_PRBC+ POSTOP_VASO+ POSTOP_HIGHCREAT+ POSTOP_COMPLICATIONS+ ACCESS_COMPLICATION+POSTOP_AH+ POSTOP_CEREBROSX+ POSTOP_RESPIRATORY+ POSTOP_DIALYSIS+ POSTOP_ARMEMBO+ POSTOP_LEGEMBO+ POSTOP_LEGCOMPART+ POSTOP_INTISCH+ POSTOP_RENALISCH+ POSTOP_SPINAL_ISCHEMIA+ RETX_R_RTOR+ DC_STATUS+ BRANCH_POST
       | PRESENTATION, data = FBVAR,overall=F, extra.col=list(`P-value`=pvalue))

knitr::kable(table1_SOC)

```

### Volume Variables 
  
Volume Variables: `REGIONID`, `CENTERID`, `PHYSICIANID`

```{r}
## ------------- clustering variables----------

#FBVAR %>% select(REGIONID) %>% table() 
#FBVAR %>% select(CENTERID) %>% table() 
#FBVAR %>% select(PHYSICIANID) %>% table() 

## ------------- plots of volume----------
center_vol = as.data.frame(FBVAR %>% select(CENTERID) %>% table())
phys_vol = as.data.frame(FBVAR %>% select(PHYSICIANID) %>% table())

p1 = ggplot(data = center_vol, aes(x=CENTERID, y=Freq)) + 
  geom_point() +
  labs(title = 'Volume of centers',x='Center ID',y='Volume')

p2 = ggplot(data = center_vol, aes(x='', y=Freq)) + 
  geom_boxplot() +
  labs(title = 'Boxplot of center volumes',x='',y='Volume')

print(ggarrange(p1, p2, widths = c(20,10),ncol = 2, nrow = 1, align = "h"))

p3 = ggplot(data = phys_vol, aes(x=PHYSICIANID, y=Freq)) + 
  geom_point() +
  labs(title = 'Volume of physicians',x='Physician ID',y='Volume')

p4 = ggplot(data = phys_vol, aes(x='', y=Freq)) + 
  geom_boxplot() +
  labs(title = 'Boxplot of physicians volumes',x='',y='Volume')

print(ggarrange(p3, p4, widths = c(20,10),ncol = 2, nrow = 1, align = "h"))

```

`r nlevels(factor(FBVAR$REGIONID))` regions, `r nlevels(factor(FBVAR$CENTERID))` centers, `r nlevels(factor(FBVAR$PHYSICIANID))` physicians. 

Quantiles of centers' volume: `r quantile(center_vol[order(center_vol$Freq),]$Freq)`

Quantiles of physicians' volume: `r quantile(phys_vol[order(phys_vol$Freq),]$Freq)`


\newpage
## Code Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,include=TRUE}
```


